---
title: "Assignment3"
author: "Marina Ramírez Baños, Giovanni Vincenzo Cavallo, Pablo Bondia Portoles, Priyanka Anantha Narayanan"
date: "2024-11-05"
output: html_document
---
# Brief explanation of our idea 

Our idea is to merge to different datasets together to find out if there is correlation between corruption in countries and the amount of non-religious people in those same countries. 
One dataset has demographic information about religion in different countries; the other dataset contains information about the Corruption Perceived Index (CPI) in those countries. 
 *Greater* values of CPI correspond to a less perceived corruption in the country.
 


# Import data

```{r}
# Load tidyverse and readr packages

library(tidyverse)
library(readr)

# Import religion_by_country and corruption_by_country datasets

religion_by_country <- read_csv("religion-by-country-2024.csv")

corruption_by_country <- read_delim("CPI2020_for_merge.csv", delim = ";", escape_double = FALSE, trim_ws = TRUE)

```

We must filter the "religion_by_country" dataset to include only the data from 2020, as is the data we have in the "corruption_by_country" dataset.


```{r}
# We have missing data for 2020 for United State, so we are gonna use the data from 2010 as an alternative

religion_by_country[religion_by_country$country == "United States", c("ReligionByCountry_TotalReligious_2020",  "religionByCountry_christians_2020", 
"religionByCountry_muslims_2020", "religionByCountry_hindus_2020", 
"religionByCountry_buddhists_2020", "religionByCountry_folkReligions_2020", 
"religionByCountry_jews_2020", "religionByCountry_other_2020", 
"religionByCountry_unaffiliated_2020")] <- religion_by_country[religion_by_country$country == "United States", c("total", "religionByCountry_christians", 
"religionByCountry_muslims", "religionByCountry_hindus", 
"religionByCountry_buddhists", "religionByCountry_folkReligions", 
"religionByCountry_jews", "religionByCountry_other", 
"religionByCountry_unaffiliated")]

#This piece of code copies the data for United States from 2010 into 2020 columns
```


```{r}

dput(names(religion_by_country))


# Filter religion_by_country dataset only including the data from 2020

religion_by_country <- religion_by_country %>% select("country", "ReligionByCountry_TotalReligious_2020",  "religionByCountry_christians_2020", 
"religionByCountry_muslims_2020", "religionByCountry_hindus_2020", 
"religionByCountry_buddhists_2020", "religionByCountry_folkReligions_2020", 
"religionByCountry_jews_2020", "religionByCountry_other_2020", 
"religionByCountry_unaffiliated_2020")



```

We create a dataset "percentage_religion_by_country", which includes the percentage of people of each religion over the total population of the country.

```{r}
# Create new columns

#Total population
percentage_religion_by_country <- religion_by_country %>% mutate(
  total_population = ReligionByCountry_TotalReligious_2020 + religionByCountry_unaffiliated_2020
)

#Proportion of individuals from each religion
percentage_religion_by_country <- percentage_religion_by_country %>% mutate(
  percentage_ReligionByCountry_TotalReligious_2020 = ((ReligionByCountry_TotalReligious_2020
)/total_population)*100,
  percentage_religionByCountry_christians_2020 = ((religionByCountry_christians_2020
)/total_population)*100,
percentage_religionByCountry_muslims_2020 =((religionByCountry_muslims_2020
)/total_population)*100,
percentage_religionByCountry_hindus_2020 = ((religionByCountry_hindus_2020
)/total_population)*100,
percetage_religionByCountry_buddhists_2020 = ((religionByCountry_buddhists_2020
)/total_population)*100,
percentage_religionByCountry_folkReligions_2020 = ((religionByCountry_folkReligions_2020
)/total_population)*100,
percentage_religionByCountry_jews_2020 = ((religionByCountry_jews_2020
)/total_population)*100,
percentage_religionByCountry_other_2020 = ((religionByCountry_other_2020
)/total_population)*100,
percentage_religionByCountry_unaffiliated_2020 = ((religionByCountry_unaffiliated_2020
)/total_population)*100
)

# Remove the other columns that do not include the proportions of individuals
percentage_religion_by_country <- percentage_religion_by_country %>% select(-starts_with("religion", ignore.case = TRUE))

```

# Merging datasets

To merge both datasets we have to use a variable they have in common. In our case, that variable is "country". However, we firstly need to rename "Country" in the corruption_by_country dataset to "country", as it is in the percentage_religion_by_country dataset.

```{r}
# Rename "Country" to "country" from corruption_by_country dataset

corruption_by_country <- corruption_by_country %>% rename(country = Country)

# Use full_join() from tidyverse to merge the datasets by country. By doing that, if a country is included in the dataset 1 but not in dataset 2, the columms of dataset 2 will contain "NA" values for that rows and viceversa

merged_dataset <- full_join(percentage_religion_by_country, corruption_by_country, by = "country")

```

# Tidying data

For the purpose of the analysis, we want to exclude those countries that do not show data about the amount of religious population in 2020 or have a NA value for CPI score 2020.


```{r}
# Remove from merged_dataset those countries with NA values for total_population

merged_dataset <- merged_dataset %>% filter(!is.na(total_population))

# Remove from merged_dataset those countries with NA values for CPI score 2020

merged_dataset <- merged_dataset %>% filter(!is.na(`CPI score 2020`))
```


## Trying to remove outliers
```{r}

# Example using IQR to remove outliers
Q1 <- quantile(merged_dataset$percentage_religionByCountry_unaffiliated_2020, 0.25, na.rm = TRUE)
Q3 <- quantile(merged_dataset$percentage_religionByCountry_unaffiliated_2020, 0.75, na.rm = TRUE)
IQR_value <- Q3 - Q1

# Filtering dataset to remove outliers
merged_dataset_filtered <- merged_dataset %>%
  filter(
    percentage_religionByCountry_unaffiliated_2020 > (Q1 - 1.5 * IQR_value) &
    percentage_religionByCountry_unaffiliated_2020 < (Q3 + 1.5 * IQR_value)
  )

```

## Applying Spearman's correlation
```{r}

#CPI Score 2020 is a chr datatype for some reason, so I try to convert to numeric
merged_dataset_filtered$`CPI score 2020` <- as.numeric(merged_dataset_filtered$`CPI score 2020`)

#Applying Spearman's correlation
cor(merged_dataset_filtered$percentage_religionByCountry_unaffiliated_2020, merged_dataset_filtered$`CPI score 2020`, method = "spearman")

```


The value of 0.3877 shows that there is a moderate positive association between CPI Score 2020 and Percentage of non-religious people in the country. This means that there is a correlation, but the association isn't that strong.

## Setting up a multiple regression model

```{r}

# Assuming merged_dataset_filtered contains the necessary variables

# Multiple Regression Model
model <- lm(`CPI score 2020` ~ 
                                percentage_religionByCountry_unaffiliated_2020, 
                                data = merged_dataset_filtered)

# Summarize the model to check coefficients and significance
summary(model)

```

By performing this linear regression, we prove that despite there being a correlation, it is not by chance and the association is statistically significant.

## Building a histogram to show Spearman's correlation
```{r}

library(ggplot2)

# 1. Histogram of Residuals
# Assuming you have the residuals from the simple regression model
simple_model <- lm(`CPI score 2020` ~ percentage_religionByCountry_unaffiliated_2020, data = merged_dataset_filtered)
merged_dataset_filtered$residuals <- residuals(simple_model)

ggplot(merged_dataset_filtered, aes(x = residuals)) +
  geom_histogram(binwidth = 2, fill = "skyblue", color = "black") +
  labs(title = "Histogram of Residuals",
       x = "Residuals",
       y = "Frequency") +
  theme_minimal()

# 2. Scatterplot with Smoothing Line
ggplot(merged_dataset_filtered, aes(x = percentage_religionByCountry_unaffiliated_2020, y = `CPI score 2020`)) +
  geom_point() +
  geom_smooth(method = "loess", color = "blue", se = FALSE) +  # Loess smoothing line for monotonic trend
  labs(title = "Scatterplot: CPI vs % Non-religious (Unaffiliated)",
       x = "% Non-religious (Unaffiliated) 2020",
       y = "CPI Score 2020") +
  theme_minimal()

```

## To build a model robust to outliers and deal with heteroscedasticity.

```{r}

#Ridge regression

x <- as.matrix(merged_dataset_filtered[, "percentage_religionByCountry_unaffiliated_2020"])
#add a dummy variable since Ridge Regression expects two columns
x <- cbind(1, x)  # Adding an intercept term
y <- merged_dataset_filtered$`CPI score 2020`

library(glmnet)
#Performing cross-validation
model <- cv.glmnet(x, y, alpha = 0)

optimal_lambda <- model$lambda.min
coef(model, s = "lambda.min")

```

This shows that for every 1% increase in the population of non-religious people, the CPI Score is expected to increase by 1.25 points. This shows that the correlation is statistically significant.

## Predict the model and visualize the relationship

```{r}


# Predicted values from the model
predictions <- predict(model, newx = x, s = "lambda.min")

# Plot the relationship
plot(merged_dataset_filtered$percentage_religionByCountry_unaffiliated_2020, predictions, 
     main = "Ridge Regression: CPI vs Percentage of Non-Religious People",
     xlab = "Percentage of Non-Religious People",
     ylab = "Predicted CPI",
     col = "blue", pch = 16)

# Add regression line
abline(lm(predictions ~ merged_dataset_filtered$percentage_religionByCountry_unaffiliated_2020), col = "red")

```