---
title: "Religion vs corruption dataset"
author: "Marina Ramírez Baños, Giovanni Vincenzo Cavallo, Pablo Bondia Portoles, Priyanka"
date: "2024-11-05"
output: html_document
---

# Import data

First, lets load the packages needed and import both data sets.

```{r}
# Load tidyverse and readr packages

library(tidyverse)
library(readr)

# Import religion_by_country and corruption_by_country datasets

religion_by_country <- read_csv("religion-by-country-2024.csv")

corruption_by_country <- read_delim("CPI2020.csv", delim = ";", escape_double = FALSE, trim_ws = TRUE, skip = 2)

```

Second, we are going to filter religion_by_country dataset to include only the data from 2020.

```{r}

dput(names(religion_by_country))


# Filter religion_by_country dataset only including the data from 2020

religion_by_country <- religion_by_country %>% select("country", "ReligionByCountry_TotalReligious_2020",  "religionByCountry_christians_2020", 
"religionByCountry_muslims_2020", "religionByCountry_hindus_2020", 
"religionByCountry_buddhists_2020", "religionByCountry_folkReligions_2020", 
"religionByCountry_jews_2020", "religionByCountry_other_2020", 
"religionByCountry_unaffiliated_2020")



```

We are going to create a dataset percentage_religion_by_country that includes the percentage of people of each religion over the total population of the country.

```{r}
# Create percentage_religion_by_country dataset

percentage_religion_by_country <- religion_by_country

# Create new columns

#Total population
percentage_religion_by_country <- percentage_religion_by_country %>% mutate(
  total_population = ReligionByCountry_TotalReligious_2020 + religionByCountry_unaffiliated_2020
)

#Proportion of individuals from each religion
percentage_religion_by_country <- percentage_religion_by_country %>% mutate(
  percentage_ReligionByCountry_TotalReligious_2020 = ((ReligionByCountry_TotalReligious_2020
)/total_population)*100,
  percentage_religionByCountry_christians_2020 = ((religionByCountry_christians_2020
)/total_population)*100,
percentage_religionByCountry_muslims_2020 =((religionByCountry_muslims_2020
)/total_population)*100,
percentage_religionByCountry_hindus_2020 = ((religionByCountry_hindus_2020
)/total_population)*100,
percetage_religionByCountry_buddhists_2020 = ((religionByCountry_buddhists_2020
)/total_population)*100,
percentage_religionByCountry_folkReligions_2020 = ((religionByCountry_folkReligions_2020
)/total_population)*100,
percentage_religionByCountry_jews_2020 = ((religionByCountry_jews_2020
)/total_population)*100,
percentage_religionByCountry_other_2020 = ((religionByCountry_other_2020
)/total_population)*100,
percentage_religionByCountry_unaffiliated_2020 = ((religionByCountry_unaffiliated_2020
)/total_population)*100
)

# Remove the other columns that do not include the proportions of individuals
percentage_religion_by_country <- percentage_religion_by_country %>% select(-starts_with("religion", ignore.case = TRUE))

```

# Merging datasets

To merge both datasets we have to use a variable they have in common. In our case, that variable is "country". However, we firstly need to rename "Country" in the corruption_by_country dataset to "country", as it is in the percentage_religion_by_country dataset.

```{r}
# Rename "Country" to "country" from corruption_by_country dataset

corruption_by_country <- corruption_by_country %>% rename(country = Country)

# Use full_join() from tidyverse to merge the datasets by country. By doing that, if a country is included in the dataset 1 but not in dataset 2, the columms of dataset 2 will contain "NA" values for that rows and viceversa

merged_dataset <- full_join(percentage_religion_by_country, corruption_by_country, by = "country")

```

# Tidying data

For the purpose of the analysis, we want to exclude those countries that do not show data about the amount of religious population in 2020 or have a NA value for CPI score 2020.

```{r}
# Remove from merged_dataset those countries with NA values for total_population

merged_dataset <- merged_dataset %>% filter(!is.na(total_population))

# Remove from merged_dataset those countries with NA values for CPI score 2020

merged_dataset <- merged_dataset %>% filter(!is.na(`CPI score 2020`))

```



Before performing the correlation analysis we check assumptions:

1. Linearity check: Pearson’s correlation measures only linear relationships. If the relationship is non-linear (e.g., quadratic or logarithmic), Pearson’s correlation may underestimate or misrepresent the true relationship.

2. 1.3. Homoscedasticity (Equal Variances):  If the variability of one variable increases as the other variable increases (called heteroscedasticity), Pearson’s correlation may not be valid, leading to incorrect conclusions about the strength of the relationship.



```{r}
# Load ggplot2 library
library(ggplot2)

# Example: Scatterplot to check linearity
ggplot(merged_dataset, aes(x = percentage_ReligionByCountry_TotalReligious_2020, y = `CPI score 2020`)) +
  geom_point() +                          # Plot the data points
  geom_smooth(method = "lm", se = FALSE, col = "red") +  # Add a linear regression line
  labs(title = "Scatterplot: Corruption vs % Affiliated",
       x = "% of affiliation", y = "% CPI") +
  theme_minimal()


```

2. Normality of variables: Pearson’s correlation assumes that the data comes from normal distributions, so violations of normality can affect the p-value and the confidence interval of the correlation coefficient, making the test less reliable. 

If it is not normal we would still do the analysis as the sample size (number of countries) is quite large so under the CLT we can make the anlaysis.

```{r}
# Histogram for each variable

# Apply log transformation
merged_dataset$Religious_Log <- log(merged_dataset$percentage_ReligionByCountry_TotalReligious_2020)

# Check histogram for CPI score 2020
ggplot(merged_dataset, aes(x = `CPI score 2020`)) + 
  geom_histogram(bins = 20, fill = "lightblue", color = "black") +
  labs(title = "Histogram of Corruption Index")

# Check histogram for original religious percentage
ggplot(merged_dataset, aes(x = Religious_Log)) + 
  geom_histogram(bins = 20, fill = "lightblue", color = "black") +
  labs(title = "Histogram of Religious Percentage")

# Q-Q plot for CPI score 2020
qqnorm(merged_dataset$`CPI score 2020`)
qqline(merged_dataset$`CPI score 2020`, col = "red")

# Q-Q plot for original religious percentage
qqnorm(merged_dataset$Religious_Log)
qqline(merged_dataset$Religious_Log, col = "red")

```

After log10 the CPI, we get a more normal like histogram although a bit skewed.


Another Check for Homoscedasticity 
```{r}

model <- lm(`CPI score 2020` ~ `Religious_Log`, data = merged_dataset)
plot(model$residuals ~ model$fitted.values)
abline(h = 0, col = "red")

```
Good Homoscedasticity 

Check for outliers:

```{r}
# Boxplot to check for outliers in 'percentage_ReligionByCountry_TotalReligious_2020'
boxplot(merged_dataset$Religious_Log,
        main = "Boxplot for Religious Percentage",
        ylab = "Religious Percentage",
        col = "lightblue")

```

Several outliers are found, however since they represent relevant cases we decide to keep them.


I tried log and sqrt transformation, the histogram looks more normal but there is still a lot of outliers and skewness, therefore a non-parametric test is suggested.
