---
title: "Religion vs corruption dataset"
author: "Marina Ramírez Baños, Giovanni Vincenzo Cavallo, Pablo Bondia Portoles, Priyanka"
date: "2024-11-05"
output: html_document
---

# Import data

First, lets load the packages needed and import both data sets.

```{r}
# Load tidyverse and readr packages

library(tidyverse)
library(readr)

# Import religion_by_country and corruption_by_country datasets

religion_by_country <- read_csv("religion-by-country-2024.csv")
corruption_by_country <- read_delim("CPI2020.csv", delim = ";", escape_double = FALSE, trim_ws = TRUE, skip = 2)

```

Second, we are going to filter religion_by_country dataset to include only the data from 2020.

```{r}

dput(names(religion_by_country))


# Filter religion_by_country dataset only including the data from 2020

religion_by_country <- religion_by_country %>% select("country", "ReligionByCountry_TotalReligious_2020",  "religionByCountry_christians_2020", 
"religionByCountry_muslims_2020", "religionByCountry_hindus_2020", 
"religionByCountry_buddhists_2020", "religionByCountry_folkReligions_2020", 
"religionByCountry_jews_2020", "religionByCountry_other_2020", 
"religionByCountry_unaffiliated_2020")



```

We are going to create a dataset percentage_religion_by_country that includes the percentage of people of each religion over the total population of the country.

```{r}
# Create percentage_religion_by_country dataset

percentage_religion_by_country <- religion_by_country

# Create new columns

#Total population
percentage_religion_by_country <- percentage_religion_by_country %>% mutate(
  total_population = ReligionByCountry_TotalReligious_2020 + religionByCountry_unaffiliated_2020
)

#Proportion of individuals from each religion
percentage_religion_by_country <- percentage_religion_by_country %>% mutate(
  percentage_ReligionByCountry_TotalReligious_2020 = ((ReligionByCountry_TotalReligious_2020
)/total_population)*100,
  percentage_religionByCountry_christians_2020 = ((religionByCountry_christians_2020
)/total_population)*100,
percentage_religionByCountry_muslims_2020 =((religionByCountry_muslims_2020
)/total_population)*100,
percentage_religionByCountry_hindus_2020 = ((religionByCountry_hindus_2020
)/total_population)*100,
percetage_religionByCountry_buddhists_2020 = ((religionByCountry_buddhists_2020
)/total_population)*100,
percentage_religionByCountry_folkReligions_2020 = ((religionByCountry_folkReligions_2020
)/total_population)*100,
percentage_religionByCountry_jews_2020 = ((religionByCountry_jews_2020
)/total_population)*100,
percentage_religionByCountry_other_2020 = ((religionByCountry_other_2020
)/total_population)*100,
percentage_religionByCountry_unaffiliated_2020 = ((religionByCountry_unaffiliated_2020
)/total_population)*100
)

# Remove the other columns that do not include the proportions of individuals
percentage_religion_by_country <- percentage_religion_by_country %>% select(-starts_with("religion", ignore.case = TRUE))

```

# Merging datasets

To merge both datasets we have to use a variable they have in common. In our case, that variable is "country". However, we firstly need to rename "Country" in the corruption_by_country dataset to "country", as it is in the percentage_religion_by_country dataset.

```{r}
# Rename "Country" to "country" from corruption_by_country dataset

corruption_by_country <- corruption_by_country %>% rename(country = Country)

# Use full_join() from tidyverse to merge the datasets by country. By doing that, if a country is included in the dataset 1 but not in dataset 2, the columms of dataset 2 will contain "NA" values for that rows and viceversa

merged_dataset <- full_join(percentage_religion_by_country, corruption_by_country, by = "country")

```

# Tidying data

For the propouse of the analysis, we want to exclude those countries that do not show data about the amount of religious population in 2020 or have a NA value for CPI score 2020.

```{r}
# Remove from merged_dataset those countries with NA values for total_population

merged_dataset <- merged_dataset %>% filter(!is.na(total_population))

# Remove from merged_dataset those countries with NA values for CPI score 2020

merged_dataset <- merged_dataset %>% filter(!is.na(`CPI score 2020`))
```

...depending on what we agree to do, continue tidying the data here
